---
title: "Assignment 3 PCA"
author: "Nicholas Jacob and Yechang Qi"
date: "2024-09-05"
output: pdf_document
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(corrplot)
library(ggfortify)
```

## Glass Data

1.a.i I'll load the data directly from my machine.  I wanted to load it directly from the web but seems as though those links don't exist anymore.  I've added the names from the info included in the zip file with the data.

```{r}
names <- c('Id', 'RI','Na','Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'Type')
Glass <- read.csv('glass.data',header = FALSE)
names(Glass)<- names
Glass$Type = as.character(Glass$Type)
head(Glass)
```

Let's double check that there isn't a duplicate since there was a warning about a possible dup.
```{r}
any(!duplicated(Glass))
```

Does not appear to be any duplicates in this dataset.

```{r}
corMat <- cor(Glass[,-11])
corMat
```
1.a.ii. Compute the eigen values and vectors of this matrix.

```{r}
eMat <- eigen(corMat)
eMat
```

As expected for the correlation matrix, all values are positive.  Eigen vectors don't show me much yet.

1.a.iii.  Next we'll do the principle component decomposition

```{r}
pMat <- prcomp(Glass[,-11],scale = TRUE)
pMat
```

1.a.iv. These are different but I don't recall why...  Look me up before you submit.

1.a.v.  Show that these vectors are orthogonal to each other.

First, I needed to see how to access the vectors
```{r}
eMat$vectors[,1] #
```

We did not find a function for doing the inner product but since the formula is 
$$
x\cdot y = \sum_{i=1}^nx_iy_i
$$
We simply asked R to do that with base packages.
```{r}
sum(pMat$rotation[,1]*eMat$vectors[,1])
```

We see that this one is indeed 1 as expected.  Let's build a matrix of all $10^2$ possible combinations.  Yes some are repeated but that does not increase runtime significantly for this data.
```{r}
idMat <- data.frame(matrix(nrow = 10, ncol = 10)) #initialize matrix
for(i in 1:10){
  for (j in 1:10){
    idMat[i,j] <- sum(pMat$rotation[,i]*eMat$vectors[,j]) #add data to matrix
  }
}
idMat #print matrix
```

We this is essentially an identity matrix but it has the overflow errors.  Let's force those to be zero.

```{r}
idMat[abs(idMat)<0.00000001] = 0 #force really small values to be zero so you can inspect
idMat
```

We might be worried about the negative ones, but we recall that the inner product is negative one if the vectors point opposite each other but otherwise the same direction.  So we do see that these two matrix have a orthonormal column space.

1.b.i.

```{r}
heatmap(corMat, main = "Heatmap of Correlation for Glass Data", scale = "column", Rowv = NA, Colv = NA)
```

I tried for a while to get a scale for the colors on this graph and failed to find it.  I am going to use the package suggest of `corrplot` and of course it works great and has the scale that I wanted in the defaults.  Yeah open source for the win!
```{r}
corrplot(corMat)
```

Yes there are two solutions here but I find it important to show that sometimes the base package just doesn't have what you want.

1.b.ii.

To look at the PCA, I'll start with the overloaded `plot` from the S3 method included with the `prcomp` function.  The scree plot follows
```{r}
plot(pMat, main = "Scree Plot for PCA of Glass", xlab = "PCA Components")
```

I don't quite see the hockey stick I expected.  I also expected this to be normalized to one but here we see variances above 1 which is odd and not what I recalled in scree plots.

```{r}
autoplot(pMat, data = Glass, color = "Type", loadings = TRUE)+
  scale_color_manual(breaks = c("1","2","3","5","6","7"),
                    values =c("red","blue","green","violet","yellow","black"))
```
```{r}

```
