---
title: "Assignment 2"
author: "Nicholas Jacob"
date: "2024-08-28"
output: pdf_document
---

```{r setup, include=TRUE}
library(tidyverse)
library(GGally)
library(ggridges)
library(naniar)
```

## 1 Learning `ggplot1`

a. \S 3.2.4 \#4 Make a scatterplot of hwy vs cyl.

```{r}
ggplot(data = mpg, aes(x = hwy,y = cyl)) +
  geom_point() +
  ggtitle("Scatter Plot of hwy VS cyl")

```
a.  \S 3.2.4 \#5 What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
```{r}
ggplot(data = mpg, aes(x = class,y = drv)) +
  geom_point() +
  ggtitle("Scatter Plot of class VS drv")

```

Since both are categorical variables, all data points are printed on top of one another.  You can use `jitter` to improve the graphic.

```{r}
ggplot(data = mpg, aes(x = class,y = drv)) +
  geom_jitter(width = 0.25) +
  ggtitle("Scatter Plot of class VS drv Now with Jitter")

```
I did relax the jitter a bit to make certain you could tell where the data belonged.

a. \S 3.3.1 \#3 Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = displ))+ 
  ggtitle("Color and Displacement added to Scatter")
```

Color gave a gradient and size made some differentiation of the size in the point.  `shape` through an error so I removed it so that I could have a picture.  The error specifically says that shape cannot be a continuous variable.

a. \S 3.3.1 \#4 What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = displ))+
  ggtitle("Adding size as one of variables")
```

Nothing, it makes the graphic that you asked for...

a. \S 3.3.1 \#6 What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = displ<5))+
  ggtitle("Including a Boolean")
```

It converted that variable from the dataset into a boolean and then graphed it.

a. \S 3.5.1 \#4 Take the first faceted plot in this section:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)+
  ggtitle("Facet Wrap of Car Type with Milage and Displacement")
```
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

Facet is going to give each `class` by itself.  You can quickly see each class and recognize where it congregates in the data.  If we had done this with the color aesthetic as below,
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))+
  ggtitle("Making Class the Color")
```

We could observe something similar in the clustering.  I am color blind so if there are too many colors, I cannot distinguish them all.  The defaults are normally fine for me but the fancier folks try to get the harder it is for me to distinguish the colors.  In a large dataset, faceting would be necessary as the color would blob over each other with some entries right on top of each other. 

b.
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy ))+
  geom_jitter(alpha = 0.5) + #jitter but not too much
  facet_wrap(~drv) + #facit into three graphs
  geom_smooth(method = "lm", se = FALSE, color = 'black' ) + #straight line of best fit
  geom_smooth(method = "loess") +  #curvy one
  ylab("Highway MPG") + #labels
  xlab("Displacement") + #labels
  ggtitle("Recreating the Master Hadley Wickham") #credit where credit is due
```
## 2. House Price Data:  EDA and Viz

```{r}
house <- read.csv('housingData-1.csv')
head(house)
```

First visualization I want to create is the kitchen sink.  I've seen this data set before so I am just going to try and get everything I can into one visualization.  I am thinking I can get 6 parts of the data...

```{r}
ggplot(data = house, aes(y = SalePrice, x = GrLivArea, color = OverallQual, size = Exterior1st, shape = RoofStyle))+
  facet_wrap(~Neighborhood) +
  geom_point()+
  scale_x_log10() +
  scale_y_log10() +
  ggtitle("Kitchen Sink Scatter")
```

Not as good as I hoped for the kitchen sink method.  Did a bit of playing around to try to make it usable but not sure you can glean much from it. I did get 6 peices of info represented in one graph so there is something to say for that.  I do like the different roof styles as those shapes too. 

Let's try another!  I did not know what was meant by sploms, but I had seen these before and even had used the `GGally` package before.  I picked the CentralAir for color because I am sweating in my office right now...
```{r}
names(house)[c(5,14,16,17,38,74,73,66)]
```
```{r}

ggpairs(house[,names(house)[c(5,14,16,17,38,74,73,66)]], aes(colour = CentralAir, alpha = 0.4))+
  ggtitle("Splom Graph of a Few Variables")
```
After a bunch of tries to get this code to compile, I limited the data available to it in the hopes of improving runtime.  I think the graphic is niffty and perhaps reveals some correlations.  Honestly, I have a hard time gleaning much from the multitude of graphics.  You have them all but my eyes glaze over and I cannot really look at any one of them for info.  Not the biggest fan of this method.

I'm going to keep moving through the suggested plots and try parallel histograms.  I am most interested in location, location, location.  So I think that effects house price the most.

```{r}
ggplot(data = house, aes(x = SalePrice, color = CentralAir))+
  geom_histogram(fill = "white") +
  facet_wrap(~Neighborhood ) +
  ggtitle("Location, Location, Location")
```
Still a bit fixated on the central air question.  Cannot really see many houses without it but where you can they mostly fall in the cheaper range.  We do see a variation in the histograms based on location.

I'll try the ridge plot next because I think those are cool
```{r}

ggplot(data = house, aes(x = SalePrice, y = Neighborhood, fill = Neighborhood))+
  geom_density_ridges()+
  theme_ridges()+ 
  theme(legend.position = "none")+
  ggtitle("Location Ridge Plot")

```
My ridge plots reinforce my previous statement that location matters.  I think this does a much better job than the histograms too.

I had never seen a missing value visualization so I'll try a few here for my last exercise.  Of course you need another library for it.

```{r}
vis_miss(house)
```

This graphic feels overwhelming.  We do see that there is a lot of missing data in a couple of columns but I am not clear which ones they are based on the names overwriting each other.  I would not have included this graphic except that I wanted to find the command again later.

Let's look for patterns in this data.
```{r}
gg_miss_upset(house)
```

This one is very informative, it shows some of the counts and the connections with missed data.  Very cool!

So I have 6 total but I find the penultimate to be cheeky and not a very good representation of the data so I did not want to include it as one of the good ones...
