---
title: "Assignment 4 Data Wrangling"
author: "Nicholas Jacob and Zayne Mclaughlin"
date: "2024-09-05"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(corrplot)
library(ggfortify)
library(MASS)
library(caret)
library(magrittr)
library(dplyr)
library(HSAUR2)
library(outliers)
library(ggbiplot)
library(GGally)
library(gridExtra)
```

1.
```{r}
housingData = read.csv("GitHub/IDA/Assignments/Assignment2/housingData-1.csv")
```


2.a.

I notice two examples of skewed data, `YearBuilt` and `SalePrice`.  The skews are in opposite directions so that is fun!  First I try the Year built.  Here is the visualization.

```{r}
yb <- ggplot(data = housingData, aes(x = YearBuilt)) +
  geom_histogram() +
  ggtitle("Year Built is Skewed") +
  xlab("Year Built")
yb
```

Very bimodal and skewed to the current time.  Apply the `boxcox` function.

```{r}
b<-boxcox(lm(housingData$YearBuilt~1), lambda = seq(-4,50,1/10)) #tweaked the limits until found the correct parameter for this
lambda <- b$x[which.max(b$y)] #find the max
lambda
```

This is the optimal $\lambda$.

```{r}
mytitle <- paste("Boxcox with lambda=", lambda)
ybm <- housingData %>%
  mutate(YearBuiltMod = (YearBuilt^lambda-1)/lambda) %>%
  ggplot( aes(x = YearBuiltMod)) +
  geom_histogram() +
  ggtitle(mytitle) +
  xlab("Year Built")

grid.arrange(yb,ybm, nrow = 1)
```
This is much better but by no means perfect.  Looks like the year built was quite a difficult column due to the bi-modal distribution.

I am going to look at the Sale Price.  These are notorious for not being normal.

```{r}
sp <- ggplot(data = housingData, aes(x = SalePrice)) +
  geom_histogram() +
  ggtitle("Sale Price is Skewed") +
  xlab("Sale Price")

sp
```
```{r}
b<-boxcox(lm(housingData$SalePrice~1), lambda = seq(-4,4,1/10))
lambda <- b$x[which.max(b$y)]
lambda
```

```{r}
mytitle <- paste("Boxcox with lambda=", lambda)
spm <- housingData %>%
  mutate(SalePriceMod = (SalePrice^lambda -1)/lambda) %>%
  ggplot( aes(x = SalePriceMod)) +
  geom_histogram() +
  ggtitle(mytitle) +
  xlab("Sale Price")

grid.arrange(sp,spm, nrow = 1)
```

That looks much better.  Quite normal and ready for analysis.

2.b.i.

Looking to `LotFrontage`, we see lots of missing values.
```{r}
missing <- is.na(housingData$LotFrontage) #find the missing values
sum(missing) #give a total
```
We impute first by replacing it with the mean.

```{r}
avg <- mean(housingData$LotFrontage, na.rm = TRUE) #get the mean before imputing
housingData <- housingData %>%
  mutate(LFMean = if_else(is.na(LotFrontage), avg,LotFrontage)) #create a new column imputed with the mean
  
```

b.ii.

Now we'll impute with a linear regression.  Since we are trying to predict something about the lot, I keep only variables with information about the outside of the house.  I could not use `Alley` because it did not have enough levels to fit the linear model.  I could not use `LotShape` due to some shapes not being in the training data.  I could not use `Fence` either due to many missing values.
```{r}
names(housingData)[c(4,5,8,9)] #find the names of variables that will work 
fit <- lm(LotFrontage ~ ., data=housingData[,names(housingData)[c(4,5,8,9)]]) #do linear fit
pred <- predict(fit,housingData[,names(housingData)[c(4,5,8,9)]] ) #create predictions
housingData <- housingData%>%
  mutate(LFLM = if_else(is.na(LotFrontage),pred,LotFrontage)) #add the new column
```
