---
title: "Assignment 6 Kaggle Competition"
author: "Nicholas Jacob and Yechang Qi"
date: "2024-09-05"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
knitr::opts_chunk$set(include = FALSE) #this will make none of the code appear unless we ask it to.
library(ggplot2)
#library(corrplot)
#library(ggfortify)
library(MASS)
library(caret)
library(magrittr)
library(dplyr)
#library(HSAUR2)
#library(outliers)
#library(ggbiplot)
#library(GGally)
#library(gridExtra)
#library(mice)
library(forcats)
library(knitr)
library(car)
library(pls)
library(lars)
```
# Preparation and Modeling

First we examined the missing data. Recycling the coding from earlier in the semester, we get the following reports on data that is missing.
```{r}
train = read.csv("Train.csv")
glimpse(train)
```
```{r}
train <- train %>% 
  group_by(custId) %>%
  #mutate(totalRevenue = sum(revenue)) %>%
  mutate(targetRevenue = log(sum(revenue) +1)) %>%
  ungroup()

glimpse(train)
```
```{r}
trainNumeric <- train %>% 
  dplyr::select(where(is.numeric))

```

```{r}
trainFactor <- train %>% 
  dplyr::select(where(is.character))%>% ##|where(is.logical) )%>%
  mutate_all(na_if,"")%>%
  mutate_all(factor)
  #mutate_all(fct_na_value_to_level) 
```



```{r}
glimpse(trainFactor)
```
```{r}
trainFactor$adwordsClickInfo.isVideoAd
```
```{r}
Q1<-function(x,na.rm=TRUE) {
quantile(x,na.rm=na.rm)[2]
}
Q3<-function(x,na.rm=TRUE) {
quantile(x,na.rm=na.rm)[4]
}
```


```{r}
myNumericSummary <- function(x){
  c(length(x), n_distinct(x), sum(is.na(x)), mean(x, na.rm=TRUE),
  min(x,na.rm=TRUE), Q1(x,na.rm=TRUE), median(x,na.rm=TRUE), Q3(x,na.rm=TRUE),
  max(x,na.rm=TRUE), sd(x,na.rm=TRUE))
}
```


```{r}
numericSummary <- trainNumeric %>%
  summarise_all(myNumericSummary) #Applying the custom myNumericSummary function to all variables in the dataset. 

glimpse(numericSummary)
```

```{r}
numericSummary <-cbind(stat=c("n","unique","missing","mean","min","Q1","median",
                              "Q3","max","sd"),numericSummary)
```

```{r}
numericSummaryFinal <- numericSummary %>%
  tidyr::pivot_longer("sessionId":"targetRevenue", names_to = "variable", values_to = "value") %>%
  tidyr::pivot_wider(names_from = stat, values_from = value) %>%#Pivoting the numeric summary table to make it longer and then wider for readability, 
  dplyr::mutate(missing_pct = 100*missing/n,
         unique_pct = 100*unique/n) %>%
  dplyr::select(variable, n, missing, missing_pct, unique, unique_pct, everything())
#Adding percentage calculations for missing and unique values. 

options(digits=3)#limit the number of digits in the table
options(scipen=99)
```

```{r, include = TRUE}
numericSummaryFinal %>% kable()#display table nicely when knitted
```

```{r}
hist(train$targetRevenue, breaks = seq(0,10,1))
```
```{r}
getmodes <- function(v,type=1) {
tbl <- table(v)
m1<-which.max(tbl)
if (type==1) {
return (names(m1)) #1st mode
}
else if (type==2) {
return (if(length(names(which.max(tbl[-m1])))==0)
{"NA"}
else {names(which.max(tbl[-m1]))}
                 ) #2nd mode
}
else if (type==-1) {
return (names(which.min(tbl))) #least common mode
}
else {
stop("Invalid type selected")
}
}
getmodesCnt <- function(v,type=1) {
tbl <- table(v)
m1<-which.max(tbl)
if (type==1) {
return (max(tbl)) #1st mode freq
}
else if (type==2) {
return (max(tbl[-m1])) #2nd mode freq
}
else if (type==-1) {
return (min(tbl)) #least common freq
}
else {
stop("Invalid type selected")
}
}
```

```{r}
myCategoricalSummary <- function(x){
  c(length(x),sum(is.na(x)), n_distinct(x),
  getmodes(x,1),getmodesCnt(x,1),getmodes(x,2),getmodesCnt(x,2),getmodes(x,-1),getmodesCnt(x,-1))
}

```
```{r}
myCategoricalSummary(trainFactor$adwordsClickInfo.adNetworkType)
```
```{r}
factorSummary <- trainFactor %>%
  summarise_all(myCategoricalSummary)

factorSummary<-cbind(stat=c("n","missing","unique",
                             "1st mode", "first_mode_freq","2nd mode",
                            "second_mode_freq","least common","least common freq"),
                     factorSummary) #add titles becarefule to omit spaces for ones we need again.
glimpse(factorSummary)
```

```{r}
factorSummaryFinal <- factorSummary %>%
  tidyr::pivot_longer("date":"adwordsClickInfo.adNetworkType", names_to = "variable", values_to = "value") %>%
  tidyr::pivot_wider(names_from = stat, values_from = value) %>%
  dplyr::mutate(missing_pct = 100*as.numeric(missing)/as.numeric(n),#compute missing_pct
         unique_pct = 100*as.numeric(unique)/as.numeric(n), #unique percent
         freqRatio = as.numeric(first_mode_freq)/as.numeric(second_mode_freq)) %>% #freqRatio as defined in assignment
  dplyr::select(variable, n, missing, missing_pct, unique, unique_pct,freqRatio, everything())

options(digits=3)
options(scipen=99)
```

```{r, include=TRUE}
factorSummaryFinal %>% kable() #display nicely
```
## Quick and Dirty Clean
```{r}
trainFactor <- train %>% 
  dplyr::select(where(is.character))%>% ##|where(is.logical) )%>%
  mutate_all(na_if,"")%>%
  mutate_all(factor)%>%
  mutate_all(fct_na_value_to_level) 
trainNumeric <- train %>% 
  dplyr::select(where(is.numeric))%>%
  replace(is.na(.),0)

train <- cbind(trainNumeric,trainFactor)

trainNumeric$nonZero <- as.factor(train$targetRevenue>0)
```
```{r}
names(trainNumeric)
```
## Identifying The Zeros
```{r}
trainControl <- trainControl(method="cv", number=5)
metric <- "Accuracy"
grid <- expand.grid(.k=seq(1,20,by=1))

fit.knn <- train(nonZero~.-targetRevenue - revenue,#
                 data=trainNumeric, method="lda",
                 metric=metric ,
                 trControl=trainControl#,
                 #tuneGrid = grid
                 )
knn.k1 <- fit.knn$bestTune # keep this Initial k for testing with knn() function in next section
print(fit.knn)
```
Once the zero's have been identified, not need to predict a value for those.  Can build the regression on just the values that are non-zero.

## Creating an Output For Competition
```{r}
test = read.csv('Test.csv')

cids = unique(test$custId)
zeros = rep(0,length(cids))#I just made them all zero for a simplistic prediction
```

```{r}
output <- cbind(custId = cids, predRevenue = zeros)
write.csv(output, "sillyResult.csv",row.names = FALSE)
```
